---
title: "Assignment 7"
author: "Francis Surroca"
date: '2022-08-03'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# Add any packages you want in this chunk:
library(plotly)
```

We are going to look at the property prices for Orange County. From there we are going to try and predict the sales price by linear regression.

# Importing Data 

Bring in the data and make sure the data types are correct. If not, make the proper changes. The file is located within this project. *data/prop_prices_reduced.csv*

```{r}
prop_prices <- read.csv("data\\prop_prices_reduced.csv")
head(prop_prices)
summary(prop_prices)

prop_prices$lake_front <- prop_prices$dist_lakes
prop_prices$lake_front[prop_prices$lake_front < 100] = 1
prop_prices$lake_front[prop_prices$lake_front > 100] = 0

prop_prices$downtown <- prop_prices$dist_cbd
prop_prices$downtown[prop_prices$downtown < 5280] = 1
prop_prices$downtown[prop_prices$downtown > 5280] = 0

prop_prices$luxury = 0
prop_prices$luxury[prop_prices$bed > 4 & prop_prices$area_heated > 2500 & prop_prices$pool == 1] = 1


# prop_prices$bed <- factor(prop_prices$bed)

head(prop_prices)
summary(prop_prices)

```


# Plotting

Plot histograms for all variables. Additionally, add scatterplots for the relationships between all quantitative variables.

1) Plot of all variables
```{r}
hist(prop_prices$sale_def)
hist(prop_prices$bed)
hist(prop_prices$bath)
hist(prop_prices$area_heated)
hist(prop_prices$area)
hist(prop_prices$dist_cbd)
hist(prop_prices$dist_lakes)
hist(prop_prices$pool)
hist(prop_prices$lake_front)
hist(prop_prices$downtown)

```

2) Scatterplots for relationships between all quantitative variables

** Sale price against all quantitative variables
```{r}
plot_ly(prop_prices, y = ~sale_def, x = ~area, type = 'scatter')
plot_ly(prop_prices, y = ~sale_def, x = ~area_heated, type = 'scatter')
plot_ly(prop_prices, y = ~sale_def, x = ~bed, type = 'scatter')
plot_ly(prop_prices, y = ~sale_def, x = ~dist_cbd, type = 'scatter')
plot_ly(prop_prices, y = ~sale_def, x = ~dist_lakes, type = 'scatter')
plot_ly(prop_prices, y = ~sale_def, x = ~bath, type = 'scatter')

plot(prop_prices$area_heated, prop_prices$sale_def)
abline(lm(sale_def ~ area_heated, data=prop_prices))

plot(prop_prices$area, prop_prices$sale_def)
abline(lm(sale_def ~ area, data=prop_prices))

plot(prop_prices$bed, prop_prices$sale_def)
abline(lm(sale_def ~ bed, data=prop_prices))

plot(prop_prices$dist_cbd, prop_prices$sale_def)
abline(lm(sale_def ~ dist_cbd, data=prop_prices))

plot(prop_prices$dist_lakes, prop_prices$sale_def)
abline(lm(sale_def ~ dist_lakes, data=prop_prices))

plot(prop_prices$bath, prop_prices$sale_def)
abline(lm(sale_def ~ bath, data=prop_prices))

```

# Summary Statistics

Provide basic summary statistics for univariate analysis. Also, provide the correlation between all the quantitative variables.

```{r}
summary(prop_prices)
cor(prop_prices)

```

# Regression Analysis

Run a regression with all the variables included. Print results of the regression.

```{r}
simple_reg = glm(sale_def ~ bed + bath + area + area_heated + dist_cbd + dist_lakes + pool + lake_front + downtown + luxury, data=prop_prices)
summary(simple_reg) 
coef(simple_reg) 

```

Which of the variables tested significant at the 95% level? Looking at the results and answering outside of the chunk is sufficient.

Answer: Pool tested significant at the 95% confidence level.

## Evaluating the model

As is, are any of the Gauss-Markov assumptions violated? If so, which ones? How can you fix the issues?

Using the plot of the regression model to test for Guass-Markov violations.

The Residuals vs Fitted plot shows that there is increasing variance as price increases indicating bias which violates assumption 1. This plot also show that there is no linearity which is also a violation.

The Q-Q plot does not show a diagnol line indicating that the data is not normally distributed.


```{r}

plot(simple_reg)

```

## New Model

Based off of your findings in the previous section, make changes to the variables, the functional form, etc.

```{r}
# Try removing outliers
normalized <- subset(prop_prices, prop_prices$sale_def < 1000000)

# Use only the most significant variables in the formula
fixed_reg = glm(sale_def ~ bed + area_heated + lake_front + luxury, data=normalized)
summary(fixed_reg) 
coef(fixed_reg) 
plot(fixed_reg)
# Looking at the plots, it seems like a better fit

# Try using log of large values to even the playing field
fixed_reg2 = glm(log(sale_def) ~ bed + log(area_heated) + lake_front + luxury, data=normalized)
summary(fixed_reg2) 
coef(fixed_reg2) 
plot(fixed_reg2)

```


# Prediction

Based on the following inputs, predict the deflated sales price:

- 2 bed
- 2 bath
- area_heated = 1223
- area = 9750
- dist_cbd = 19368
- dist_lakes = 490
- no pool

```{r}
# Using basic model
new_house_df <- data.frame(bed=2,bath=2,area_heated=1223,area=9750,dist_cbd=19368,dist_lakes=490,pool=0,lake_front=0,luxury=0)
basic_linear_model <- lm(formula = sale_def ~ bed + bath + area + area_heated + dist_cbd + 
    dist_lakes + pool, data = prop_prices)
predict(basic_linear_model, newdata=new_house_df)

# Now using fixed models
predict(fixed_reg, newdata=new_house_df)

# Need to take e^predication since fixed_reg2 is based on logs
exp(predict(fixed_reg2, newdata=new_house_df))


```
